---
title: "portfolioAnalysis"
author: "Niraj Sardar"
output: html_document
---

```{r setup, include=FALSE}
library(quantmod)
```

## Mean Value analysis of portfolio

### A. Loading Ticker Data
```{r}

ticks <- c("^GSPC", "CAT", "LLY")

retout <- NULL
retout <- xts(retout)

for(i in 1:length(ticks)){
  prices <-  getSymbols(ticks[i], auto.assign = F)
  returns <- periodReturn(prices, period = "monthly", 
                          type = "arithmetic")
  retout <- merge.xts(retout, returns)
}

colnames(retout) <- ticks
retout = retout['2013/2016']
head(retout)
```

### B. Calculating mean vector and covariance matrix


```{r}
meanret <- apply(retout,2,mean)
meanret


covar <- var(retout)
covar
```

```{r}
weight <- c(0, .4, .6)

```

### C. Calculating portfolio mean and variance

```{r}

rp <- weight[2]*meanret[2] + weight[3]*meanret[3]
rp <- unname(rp) # otherwise picks up second ticker as dimname
rp

```


```{r}
sig2p <- weight[2]^2*covar[2,2]+weight[3]^2*covar[3,3]+2*weight[2]*weight[3]*covar[2,3]
sig2p

```


```{r}
sqrt(sig2p)
```


### D. Calculating using matrix multiplication

```{r}

weight <- as.matrix(weight)
dim(weight)

meanret <- as.matrix(meanret)
dim(meanret)

```

Use matrix multiplication to calculate portfolio metrics:

```{r}
rp <- t(weight) %*% meanret
rp

sig2p <- t(weight) %*% covar %*% weight
sig2p

sigp <- sqrt(sig2p)   # portfolio sigma
sigp

```


Matrix algebra is useful inorder to avoid complex for loops.

## E. Feasible Set with these two assets

Now we are ready to do multiple computations. Simulate various portfolios and calculate mean and sigma.


```{r}

# initialize a counter and the results matrix
kount <- 0
Results <- matrix(data = NA, nrow = length(seq(.05,.95,.05)), ncol = 4)

for (i in seq(.05,0.95,.05)){
    kount <- kount + 1 # counter for portfolio number
    Results[kount,1] = weight[2,1] = i        # weight of security 1
    Results[kount,2] = weight[3,1] = 1-i      # weight of security 2
    Results[kount,3] <- t(weight) %*% meanret # portfolio mean
    Results[kount,4] <- sqrt(t(weight) %*% covar %*% weight) # portfolio sigma
}
colnames(Results) <- c(paste0(ticks[2], '%'), paste0(ticks[3], '%'), "Port_Mean", "Port_Sigma")

Results

```

Plot feasible set

```{r}

plot(x = Results[,4], y = Results[,3], xlab = "Portfolio Sigma", ylab = "Portfolio Mean")

```