---
title: "Portfolio Efficient Set"
author: "Niraj Sardar"
output: html_document
---

Loading libraries

```{r}
library(quantmod)
library(ggplot2)
library(dplyr)
```


```{r}
ticks <- c("STZ", "CAT", "LLY", "CBRL")

retout <- NULL
retout <- xts(retout)

for(i in 1:length(ticks)){
  prices = getSymbols(ticks[i], auto.assign = F)
  returns <- periodReturn(prices, period = "monthly", 
                          type = "arithmetic")
  retout <- merge.xts(retout, returns)
}
colnames(retout) <- ticks
retout <- retout['2010/2015']
retout <- na.omit(retout)
head(retout)

```

### B. Calculate means and covariance matrix

```{r}
meanret <- colMeans(retout,na.rm = T)
x1 = round(meanret, 5)
cat("The mean vector :\n")
print(x1)

```


```{r}
covar <- var(retout)
x2 = round(covar, 8)
cat("The covariance matrix: \n")
print(x2)

```

### C. Creating  one portfolio as an example to showcase method

The weight vector contains weights for each security. Weights add to 1.

```{r}
weight <- c(.2, .3, .3, .2)
```

Calculate portfolio variance Using matrix algebra

```{r}
weight <- as.matrix(weight)
dim(weight)

meanret <- as.matrix(meanret)
dim(meanret)

mretp <- t(weight) %*% meanret
sretp <- sqrt(t(weight) %*% covar %*% weight)

cat("The mean and sigma of portfolio returns: ", mretp, sretp)
```

### D. Run simulation with random weights

First, for niter iterations, let’s create random portfolio weights.


```{r}
set.seed(12)
niter <- 500   # Set the number of iterations here
randomnums <- data.frame(replicate(4, runif(niter, 1, 10)))

head(randomnums)

wt_sim <- randomnums / rowSums(randomnums)
cat("The weights after normalization are in wt_sim...")

head(wt_sim)


```

Initializing Variables.

```{r}
# initialize weight and Results matrices
weight <- matrix(data = NA, nrow = length(ticks), ncol = 1)
Results <- matrix(data = NA, nrow = niter, ncol = 6)
```

Run the simulations - this means, do portfolio calculations for each simulated portfolio.

```{r}

# loop: each i is a portfolio
for (i in 1:niter){

    # inner loop places weights into Results
    for (k in 1:length(ticks)) {
             Results[i,k] = weight[k,1] = wt_sim[i,k]
    }

    Results[i,5] <- t(weight) %*% meanret                # portfolio mean
    Results[i,6] <- sqrt(t(weight) %*% covar %*% weight) # portfolio sigma
}

colnames(Results) <- c(ticks, "PortMean", "PortSigma")
Results <- as.data.frame(Results)
head(Results)

```

Plotting the results.

```{r}
ggplot(data = Results , aes(x = PortSigma, y = PortMean)) +
    geom_point(pch = 10, colour = "red", size = 3)
```

### E.  Optimization

Again, risk is bad while return is good, so one way to select the best portfolio(s) is to identify a constraint based on one of these dimensions, and locate the best portfolio using the other.

```{r}

minmret = min(Results$PortMean)
maxmret = max(Results$PortMean)
seqmret = seq(round(minmret,3)-.001, maxmret+.001, .001)

optim <- Results %>% mutate(portnumber = index(Results)) %>%
    mutate(ints = cut(PortMean ,breaks = seqmret), 
           lower = as.numeric( sub("\\((.+),.*", "\\1", ints) )) %>% 
    group_by(ints) %>% 
    summarise( lowerval = min(lower),
               sig_optim = min(PortSigma), 
               retn_optim = PortMean[which.min(PortSigma)],
               numb = length(PortSigma), 
               portID=portnumber[which.min(PortSigma)])

optim


ggplot(data = optim , aes(x = sig_optim, y = retn_optim)) +
    geom_point(pch = 10, colour = "red", size = 3)
```


```{r}

```
